Parallelizing Linear Recurrent Neural Nets Over Sequence Length

### 2017-10-27 paper/main.tex
Hid final comment from me
Fixing stuff

\begin{abstract}
Recurrent neural networks (RNNs) are widely used to model sequential data but
their non-linear dependencies between sequence elements prevent parallelizing
training over sequence length. We show the training of RNNs with only linear
sequential dependencies can be parallelized over the sequence length using the
parallel scan algorithm, leading to rapid training on long sequences even with
small minibatch size. We develop a parallel linear recurrence CUDA kernel and
show that it can be applied to immediately speed up training and inference of
several state of the art RNN architectures by up to 9x.  We abstract recent work
on linear RNNs into a new framework of linear surrogate RNNs and develop a
linear surrogate model for the long short-term memory unit, the GILR-LSTM, that
utilizies parallel linear recurrence.  We extend sequence learning to new
extremely-long-timestep regimes that were previously out of reach by
successfully training a GILR-LSTM on a synthetic sequence classification task
with a one million timestep dependency.
\end{abstract}

@@

\section{Experiments}
We perform several experiments. First we find that our parallel linear recurrence
kernel is able to able to achieve up to 40x higher throughput than a serial implementation
when applied to long sequences. Secondly, we confirm that this kernel speedup
translates to up to a 9x speedup to LS-RNNs such as QRNNs.

In order to illustrate that the linearization does not necessarily come at the
cost of expressibility, we show that the GILR-LSTM architecture computed with
the parallel linear recurrence algorithm is able to train significantly faster
than an optimized LSTM implementation on a pathological long-term dependency
problem from the original LSTM paper \citep{hochreiter1997long}.
Finally we show that the GILR-LSTM is able to
quickly achieve state of the art results on an existing medical data
set. \Chris{I hope so\ldots}

@@

We ran all experiments on a NVIDIA K80 GPU, with five runs per configuration
allowing us to find the average and standard deviation of the time and number of
iterations to convergence. We continually generated random sequences to serve
as input data.  A brief search over learning rate
and batch size was carried out to find the parameters which allow the network to
converge most rapidly for all runs. The criterion for convergence was five consecutive
minibatches giving 100\% accuracy. The learning curves
in figure \ref{fig:synthetic_training} give support to this being a reasonable
convergence criteria.
For the longest sequence length, we did not observe the CuDDN converging, even after
several days' training.
\begin{table}[]
\centering
\caption{Performance of the GILR-LSTM compared to the CuDNN-LSTM
  on problem 2b from \citet{hochreiter1997long}. We see that the GILR-LSTM has a
  clear advantage in training speed as the sequence length increases.  We did
  not observe the Cudnn-LSTM converging for the longest task, even after several
  days.}
\label{my-label}
\begin{tabular}{@{}lllllll@{}} \toprule Sequence Length &
\multicolumn{2}{c}{\textbf{1,024}} & \multicolumn{2}{c}{\textbf{8,192}} &
\multicolumn{2}{c}{\textbf{1,048,576}} \\ \midrule &CuDNN& GILR &
CuDNN &GILR&CuDNN&GILR\\ \cmidrule(l){2-7} Iterations (1000s) &
                     1.0 %bs=8
                     \(\pm\) 0.4 & 0.55 %bs=8
                                    \(\pm\) 0.04 & 0.44 %bs=8
                                                   \(\pm\) 0.05 & 0.56 %bs=8
                                                                 \(\pm\) 0.16
                                                               &-& 14 \(\pm\) 3 \\
\begin{tabular}[c]{@{}l@{}}Wall time (hours)\end{tabular} &
0.28 \(\pm\) 0.08 & 0.031 \(\pm\) 0.002 & 0.58 \(\pm\) 0.06 & 0.10 \(\pm\) 0.03
&-& 9.7 \(\pm\) 1.7 \\ \bottomrule
\end{tabular}
\end{table}

The results illustrate that the GILR-LSTM is able to converge between 6 and 10
times faster than the CuDNN-LSTM. This is somewhat surprising
given the LSTM was specifically constructed for problems of this sort, and the
CuDNN-LSTM is a highly-optimised implementation of the LSTM (to the extent that
the monolithic interface it exposes is difficult to understand or extend). The
GILR-LSTM is written almost entirely in native tensorflow, with only the call to
the linear recurrence kernel activating an outside hook. This GILR-LSTM
convergence leads to the conclusion that the non-linearities present in LSTM are
not necessary for solving this instance of the long-term dependency problem.
The time to convergence further leads to the conclusion that inclusion of a
non-linearity at every step incurs a significant training time slowdown.
Furthermore, the GILR-LSTM is able to learn to carry dependencies over a one
million element sequence. As far as we know, this one million step sequence
experiment is the longest sequential learning problem to be handled by neural
networks to date.

@@

\section{Conclusion}
A significant portion of the success of deep learning can be attributed to massive
amounts of computation. Most of this computation is accessed through two
building blocks with highly efficient parallel implementations:
matrix multiplication and convolution. Recent research has demonstrated that
linear RNNs can achieve similar prediction accuracy to non-linear RNNs on a wide
variety of tasks in a fraction of the training time. We propose the framework of
LS-RNNs as a way to tame the growing zoo of sequential neural nets. We identify linear
recurrence as another parallelizable building block for current and future sequential models
and we use it to obtain significant speedups on already fast models.
With the power of parallel
linear recurrence we are able to solve a sequential dependency problem multiple
orders of magnitude larger than anything done prior.
Future applications of parallel linear recurrence within neural nets could include
parallel training of memory augmented models or providing a new sort of image filter
on very high resolution images.
We hope that parallel linear recurrence can be to large scale sequence
modelling what fast convolution algorithms are to image recognition.

>>>>

\begin{abstract}
Recurrent neural networks (RNNs) are widely used to model sequential data but
their non-linear dependencies between sequence elements prevent parallelizing
training over sequence length. We show the training of RNNs with only linear
sequential dependencies can be parallelized over the sequence length using the
parallel scan algorithm, leading to rapid training on long sequences even with
small minibatch size. We develop a parallel linear recurrence CUDA kernel and
show that it can be applied to immediately speed up training and inference of
several state of the art RNN architectures by up to 9x.  We abstract recent work
on linear RNNs into a new framework of linear surrogate RNNs and develop a
linear surrogate model for the long short-term memory unit, the GILR-LSTM, that
utilizes parallel linear recurrence.  We extend sequence learning to new
extremely long sequence regimes that were previously out of reach by
successfully training a GILR-LSTM on a synthetic sequence classification task
with a one million timestep dependency.
\end{abstract}

@@

\section{Experiments}
We perform several experiments. First we find that our parallel linear recurrence
kernel is able to achieve up to 40x higher throughput than a serial implementation
when applied to long sequences. Secondly, we confirm that this kernel speedup
translates to up to a 9x speedup to LS-RNNs such as QRNNs.

In order to illustrate that the linearization does not necessarily come at the
cost of expressibility, we show that the GILR-LSTM architecture computed with
the parallel linear recurrence algorithm is able to train significantly faster
than an optimized LSTM implementation on a pathological long-term dependency
problem from the original LSTM paper \citep{hochreiter1997long}.

@@

We ran all experiments on a NVIDIA K80 GPU, with five runs per configuration
allowing us to find the average and standard deviation of the time and number of
iterations to convergence. We continually generated random sequences to serve
as input data.  A brief search over learning rate
and batch size was carried out to find the parameters which allow the network to
converge most rapidly for all runs. The criterion for convergence was five consecutive
minibatches giving 100\% accuracy. The learning curves
in figure \ref{fig:synthetic_training} give support to this being a reasonable
convergence criteria.
For the longest sequence length, we did not observe the CuDDN converging, even after
several days' training.
\begin{table}[]
\centering
\caption{Performance of the GILR-LSTM compared to the CuDNN-LSTM
  on problem 2b from \citet{hochreiter1997long}. We see that the GILR-LSTM has a
  clear advantage in training speed as the sequence length increases.  We did
  not observe the Cudnn-LSTM converging for the longest task, even after several
  days.}
\label{table:synth-table}
\begin{tabular}{@{}lllllll@{}} \toprule Sequence Length &
\multicolumn{2}{c}{\textbf{1,024}} & \multicolumn{2}{c}{\textbf{8,192}} &
\multicolumn{2}{c}{\textbf{1,048,576}} \\ \midrule &CuDNN& GILR &
CuDNN &GILR&CuDNN&GILR\\ \cmidrule(l){2-7} Iterations (1000s) &
                     1.0 %bs=8
                     \(\pm\) 0.4 & 0.55 %bs=8
                                    \(\pm\) 0.04 & 0.44 %bs=8
                                                   \(\pm\) 0.05 & 0.56 %bs=8
                                                                 \(\pm\) 0.16
                                                               &-& 14 \(\pm\) 3 \\
\begin{tabular}[c]{@{}l@{}}Wall time (hours)\end{tabular} &
0.28 \(\pm\) 0.08 & 0.031 \(\pm\) 0.002 & 0.58 \(\pm\) 0.06 & 0.10 \(\pm\) 0.03
&-& 9.7 \(\pm\) 1.7 \\ \bottomrule
\end{tabular}
\end{table}

The results as show in table \ref{synth-table} illustrate that the GILR-LSTM is able to converge between 6 and 10
times faster than the CuDNN-LSTM. This is somewhat surprising
given the LSTM was specifically constructed for problems of this sort, and the
CuDNN LSTM implementation is highly optimized (to the extent that
the monolithic interface it exposes is difficult to modify or extend). The
GILR-LSTM is implemented entirely in standard TensorFlow with the exception of
the call to new linear recurrence kernel in place of a TensorFlow symbolic
loop.
Convergence of the GILR-LSTM models
leads to the conclusion that the non-linearities present in LSTM are
not necessary for solving this instance of the long-term dependency problem.
The time to convergence further leads to the conclusion that inclusion of a
non-linearity at every step incurs a significant training time slowdown.
Furthermore, the GILR-LSTM is able to learn to carry dependencies over a one
million element sequence. As far as we know, this one million step sequence
experiment is the longest sequential learning problem to be handled by neural
networks to date.

@@

\section{Conclusion}
A significant portion of the success of deep learning can be attributed to access to massive
amounts of computation. Most of this computation is accessed through two
highly efficient and parallelizable building blocks:
matrix multiplication and convolution. Recent research has demonstrated that
linear RNNs can achieve similar prediction accuracy to non-linear RNNs on a wide
variety of tasks in a fraction of the training time. We propose the framework of
LS-RNNs as a way to tame the growing zoo of sequential neural nets. We identify linear
recurrence as another parallelizable building block for current and future sequential models
and we use it to obtain significant speedups on already fast models.
With the power of parallel
linear recurrence we are able to solve a sequential dependency problem multiple
orders of magnitude larger than anything done prior.
Future applications of parallel linear recurrence within neural nets could include
parallel training of memory augmented models or providing a new sort of image filter
on very high resolution images.
We hope that parallel linear recurrence can be to large scale sequence
modelling what fast convolution algorithms are to image recognition.
